Parallel Global Optimization for Non-Convex Mixed-Integer Problems \thanks{This study was supported by the Russian Science Foundation, project No.\,16-11-10150.}

### 2019-03-28 2019/RSD_Integer/RSD2019.tex

\title{Parallel Global Optimization for Non-Convex Mixed-Integer Problems}
\author{Konstantin Barkalov \and Ilya Lebedev %\Letter 
\\
\email{konstantin.barkalov@itmm.unn.ru}}

@@

By employing the continuous single-valued Peano curve $y(x)$ mapping the unit interval 
$[0,1]$ on the $x$-axis onto the $N$-dimensional domain (\ref{D}), it is possible to find the 
minimum in (\ref{problem}) by solving the one-dimensional problem
\[
\varphi(y(x^\ast))=\min \left\{\varphi(y(x)): x \in [0,1], \; g_i(y(x))\leq 0, \; 1 \leq i \leq m\right\}.
\]
Algorithms for numerical construction of Peano curve approximation (\textit{evolvent}) are 
considered in \cite{Strongin2013}. Due to (\ref{g_k}) the functions $g_i(y(x))$ are defined 
and computable in the subranges 
\[
Q_1=[0,1], \; Q_{i+1}=\left\{x \in Q_i : g_i(y(x)) \leq 0 \right\}, \; 1 \leq i \leq m.
\]

@@

Thus, \textit{a trial} at a point $x^k \in [0,1]$ executed at the $k$-th iteration of the algorithm 
will consist of the following sequence of operations:
\begin{itemize}
	\item Determine the \textit{image} $y^k=y(x^k)$ in accordance with the mapping 
$y(x)$;
	\item Compute the values $g_1(y^k),..., g_\nu(y^k),$ where $\nu = \nu(x^k)$ is from 
(\ref{nu}). 
\end{itemize}
The dyad $ \{ \nu=\nu(x^k), \; z^k=g_\nu(y(x^k)) \} $ will be referred to as the \textit{trial 
outcome}.

An efficient parallel index algorithm (PIA) for solving the constrained optimization problem 
(\ref{problem}) has been developed at University of Nizhni Novgorod.

@@

\begin{enumerate}
\item 
Renumber the points $x^1,...,x^k$ from previous iterations with lower indices, lowest to highest coordinate values, i.e.
\begin{equation}\label{Eq:17}
0=x_0<x_1<...<x_i<...<x_k<x_{k+1}=1,
\end{equation}
and match them with the values $z_i=g_\nu(y(x_i))$, $\nu=\nu(x_i)$, $1 \leq i \leq k$, from (\ref{Eq:16}) , calculated at these points; points $x_0=0$ и $x_{k+1}=1$ are introduced additionally; the values $z_0$ и $z_{k+1}$ are indeterminate.
\item
Classify the numbers $i,1\leq i \leq k$, of the trial points from (\ref{Eq:17}) by the number of problem constraints fulfilled at these points, by building the sets
\begin{equation}\label{Eq:18}
I_\nu = \left\{i: 1 \leq i \leq k,\ \nu = \nu(x_i)\right\},\ 1 \leq \nu \leq m+1,
\end{equation}
containing the numbers of all points $x_i,1\leq i \leq k$, with the same values of $\nu$. The end points $x_0=0$ and $x_{k+1}=1$ are interpreted as those with zero indices, and they are matched to an additional set $I_0={0,k+1}$.

@@

Рассмотрим теперь случай, когда аргумент функций задачи содержит две компоненты: вектор $y$, принадлежащий гиперинтервалу $D$, и вектор $u$, имеющий конечный (и при этом не очень большой) набор возможных значений, т.е. 
\begin{gather}\label{problem_i}
\min{\left\{ g_{m+1}(y,u):y\in D, \; g_j(y,u)\leq 0, \; 1 \leq j \leq m\right\}},\\
D=\left\{a_i\leq y_i \leq b_i, \; 1\leq i \leq N \right\}.\nonumber
\end{gather}

@@

Занумеруем целыми числами $s, 1\leq s \leq S,$ все возможные значения вектора $u$, т.е. сопоставим каждому рассматриваемому значению $s$ вектор $u_s$. 
Тогда рассматриваемая задача может быть записана в виде 
\begin{gather}\label{problem_is}
 \min_{s\in\{1,...,S\}}\left(\min{\left\{ g_{m+1}(y,u_s):y\in D, \; g_j(y,u_s)\leq 0, \; 1 \leq j \leq m\right\}}\right),\\
D=\left\{ a_i\leq y_i \leq b_i, \;  y_i \in R, 1 \leq i\leq N \right\}.\nonumber 
\end{gather}

Используя схему редукции размерности с помощью развертки $y(x), x\in [0,1]$ можно сопоставить каждой вложенной задаче минимизации по $y$ одномерную задачу минимизации
\[
 \min{\left\{ g_{m+1}(y(x),u_s):x \in [0,1], \; g_j(y(x),u_s)\leq 0, \; 1 \leq j \leq m\right\}}, s\in\{1,...,S\}.
\]

Рассмотрим теперь соответствие 
\[
Y(x)=y(x-E(x)), \; x\in[0,S],
\]
отображающее любую точку интервала [0,S] на область $D$ (обозначение $E(x)$ соответствует целой части числа $x$) и определим функции 
\[
g_j(x) = g_j(Y(x),u_{E(x)+1}), x\in[0,S],
\]
имеющие, вообще говоря, разрывы первого рода в целочисленных точках $x_i = i, 1\leq i \leq S-1$.
%значения функций и индексов в данных точках - не определены!
Используя введенные обозначения можно переформулировать исходную задачу как
\begin{equation}\label{problem_is1}
\min \left\{g_{m+1}(x): x \in [0,S], \; g_j(x) \leq 0, \; 1 \leq j \leq m\right\}.
\end{equation}

@@

\begin{figure}[ht]
    \centering
    \includegraphics[height=2.9cm,width=0.8\textwidth]{fig1.jpg}
    \caption{Reduced mixed-integer global optimization problem}
    \label{fig:1}
\end{figure}

@@

In order to demonstrate the reliability of the MIGSA method, we have solved two series of 100 
multiextremal mixed-integer problems each constructed on the basis of modified problems of 
\textit{Simple} and \textit{Hard} classes generated by GKLS generator. %~\cite{Gaviano}. 
This generator of multiextremal functions is often used for the investigations of the global optimization algorithms~ 
\cite{Paulavicius2014,SergeyevKvasov2015,Lebedev2015,Gergel2015}.
In the problems generated in our experiments, there were 5 integer  and 3 continuous parameters.
The accuracy of the search %of the solution in the coordinates 
was equal to $10^{-2}$. All  
problems of the series have been solved successfully and for solving the problems based on 
the \textit{Simple} class 11988 trials in average were required whereas for solving the problems based 
on the \textit{Hard} class 24750 trials were required.

\textbf{Acknowledgments}. This study was supported by the Russian Science Foundation, project No 16-11-10150.

@@

%\bibitem{Gaviano} Gaviano, M., Kvasov, D.E, Lera, D., and Sergeyev, Ya.D.: Software for generation of classes of test functions with known local and global minima for global optimization. ACM Transactions on Mathematical Software \textbf{29}(4), 469--480 (2003)

>>>>

\title{Parallel Global Optimization for Non-Convex Mixed-Integer Problems
\thanks{This study was supported by the Russian Science Foundation, 
project No.\,16-11-10150.}}
\author{Konstantin Barkalov \and Ilya Lebedev %\Letter 
\\
\email{konstantin.barkalov@itmm.unn.ru}}

@@

By employing the continuous single-valued Peano-Hilbert curve $y(x)$ mapping the unit interval 
$[0,1]$ on the $x$-axis onto the $N$-dimensional domain (\ref{D}), it is possible to find the 
minimum in (\ref{problem}) by solving the one-dimensional problem
\[
\varphi(y(x^\ast))=\min \left\{\varphi(y(x)): x \in [0,1], \; g_i(y(x))\leq 0, \; 1 \leq i \leq m\right\}.
\]
Algorithms for numerical construction of Hilbert curve approximation (\textit{evolvent}) are 
considered in \cite{Strongin2013}. These evolvents are fractals generated by an iterative
process, that fill in the hypercube $D$ with accuracy $2^{-m}$, where integer $m>0$ is the evolvent construction parameter. 
%Пример разверток - рисунки: 2D или 3D?
%Examples of the evolvent with different m in two dimensions are given in Fig. 1.
%In Fig. 1 approximations of levels 1, 2, 3, and 4 of the Hilbert curve are given in the domain [􀀀1; 1] \in R3.

Due to (\ref{g_k}) the functions $g_i(y(x))$ are defined 
and computable in the subranges 
\[
Q_1=[0,1], \; Q_{i+1}=\left\{x \in Q_i : g_i(y(x)) \leq 0 \right\}, \; 1 \leq i \leq m.
\]

@@

Thus, \textit{a trial} at a point $x^k \in [0,1]$ executed at the $k$-th iteration of the algorithm 
will consist of the following sequence of operations:
\begin{itemize}
	\item Determine the \textit{image} $y^k=y(x^k)$ in accordance with the mapping 
$y(x)$;
	\item Compute the values $g_1(y^k),..., g_\nu(y^k),$ where $\nu = \nu(x^k)$ is from 
(\ref{nu}). 
\end{itemize}
The dyad 
\begin{equation} \label{trial_result}
 \{ \nu=\nu(x^k), \; z^k=g_\nu(y(x^k)) \} 
\end{equation}
will be referred to as the \textit{trial outcome}.

An efficient parallel index algorithm (PIA) for solving the constrained global optimization problem (\ref{problem}) has been developed at University of Nizhni Novgorod.

@@

\begin{enumerate}
\item 
Renumber the points $x^1,...,x^k$ from previous iterations with lower indices, lowest to highest coordinate values, i.e.
\begin{equation}\label{Eq:17}
0=x_0<x_1<...<x_i<...<x_k<x_{k+1}=1,
\end{equation}
and match them with the values $z_i=g_\nu(y(x_i))$, $\nu=\nu(x_i)$, $1 \leq i \leq k$, from (\ref{trial_result}), calculated at these points; points $x_0=0$ и $x_{k+1}=1$ are introduced additionally; the values $z_0$ и $z_{k+1}$ are indeterminate.
\item
Classify the numbers $i,1\leq i \leq k$, of the trial points from (\ref{Eq:17}) by the number of problem constraints fulfilled at these points, by building the sets
\begin{equation}\label{Eq:18}
I_\nu = \left\{i: 1 \leq i \leq k,\ \nu = \nu(x_i)\right\},\ 1 \leq \nu \leq m+1,
\end{equation}
containing the numbers of all points $x_i,1\leq i \leq k$, with the same values of $\nu$. The end points $x_0=0$ and $x_{k+1}=1$ are interpreted as those with zero indices, and they are matched to an additional set $I_0={0,k+1}$.

@@

Рассмотрим теперь случай, когда аргумент функций задачи содержит две компоненты: вектор $y$, принадлежащий гиперинтервалу $D$, и вектор $u$, имеющий конечный (и при этом не очень большой) набор возможных значений, т.е. 
\begin{gather}\label{problem_i}
\min{\left\{ g_{m+1}(y,u):y\in D, \; g_i(y,u)\leq 0, \; 1 \leq i \leq m\right\}},\\
D=\left\{a_j\leq y_j \leq b_j, \; 1\leq j \leq N \right\}.\nonumber
\end{gather}

@@

Занумеруем целыми числами $s, 1\leq s \leq S,$ все возможные значения вектора $u$, т.е. сопоставим каждому рассматриваемому значению $s$ вектор $u_s$. 
Тогда рассматриваемая задача может быть записана в виде 
\begin{gather}\label{problem_is}
 \min_{s\in\{1,...,S\}}\left(\min{\left\{ g_{m+1}(y,u_s):y\in D, \; g_i(y,u_s)\leq 0, \; 1 \leq i \leq m\right\}}\right),\\
D=\left\{ a_j\leq y_j \leq b_j, \; 1 \leq j\leq N \right\}.\nonumber 
\end{gather}

Используя схему редукции размерности с помощью развертки $y(x), x\in [0,1]$ можно сопоставить каждой вложенной задаче минимизации по $y$ одномерную задачу минимизации
\[
 \min{\left\{ g_{m+1}(y(x),u_s):x \in [0,1], \; g_i(y(x),u_s)\leq 0, \; 1 \leq i \leq m\right\}}, s\in\{1,...,S\}.
\]

Рассмотрим теперь соответствие 
\[
Y(x)=y(x-E(x)), \; x\in[0,S],
\]
отображающее любую точку интервала [0,S] на область $D$ (обозначение $E(x)$ соответствует целой части числа $x$) и определим функции 
\[
g_i(x) = g_i(Y(x),u_{E(x)+1}), x\in[0,S],
\]
имеющие, вообще говоря, разрывы первого рода в целочисленных точках $x_k = i, 1\leq i \leq S-1$.
%значения функций и индексов в данных точках - не определены!
Используя введенные обозначения можно переформулировать исходную задачу как
\begin{equation}\label{problem_is1}
\min \left\{g_{m+1}(x): x \in [0,S], \; g_i(x) \leq 0, \; 1 \leq i \leq m\right\}.
\end{equation}

@@

\begin{figure}[ht]
    \centering
    \includegraphics[width=1.0\textwidth] {fig1.jpg}
    \caption{Reduced mixed-integer global optimization problem}
    \label{fig:1}
\end{figure}

@@

%In order to demonstrate the reliability of the MIGSA method, we have solved two series of 100 multiextremal mixed-integer problems each constructed on the basis of modified problems of \textit{Simple} and \textit{Hard} classes generated by GKLS generator.

Известным генератором тестовых задач непрерывной многоэкстремальной оптимизации является GKLS-генератор \cite{Gaviano}. Он позволяет генерировать functions of arbitrary dimensionality with known properties (the number of local minima, the size of their domains of attraction, the global minimizer, etc.).  
This generator of multiextremal functions is often used for the investigations of the global optimization algorithms~ 
\cite{Paulavicius2014,SergeyevKvasov2015,Lebedev2015,Gergel2015}.
%Линии уровня какой-нибудь задачи GKLS? C точками испытаний?

В проведенных экспериментах GKLS-генератор использовался как основа для построения mixed-integer problems. Правила, позволяющие генерировать тестовые задачи подобного вида, заключаются в следующем.

\begin{enumerate}
	\item С помощью генератора GKLS порождается непрерывная многоэкстремальная функция $\varphi(y), \; y\in D, \; D = \left\{ a_j\leq y_j\leq b_j, 1\leq j \leq N \right\}$. Глобальный минимум данной функции достигается в известной точке $y'=(y'_1,...,y'_N)$ и равен $\varphi'=\varphi(y')=-1$.
	\item Генерируется concave mixed-integer function 
	\[
			\mu(y,u) = -2 \left[ \sum_{j=1}^N \left( \frac{y_j - y'_j}{b_j-a_j} \right)^2 + \sum_{j=1}^M \left( \frac{u_j - b_j}{b_j-a_j} \right)^2 \right].
	\]
	где $y\in D = \left\{ a_j\leq y_j\leq b_j, 1\leq j \leq N \right\} \subset R^N$, a $u\in U = = \left\{ a_j\leq u_j\leq b_j, 1\leq j \leq M \right\} \subset Z^N$; т.е. данная функция имеет $N$ непрерывных и $M$ дискретных параметров и достигает минимального значения в точке $(y',b)$.
	\item Вычисляется коэффициент 
	\[
	C = 4 - \min_{y,u} \left\{ \mu(y,u) \right\}, \; y\in D, u \in U.
	\]
	Очевидно, что минимум вогнутой функции $\mu(y,u)$ в одной из угловых точек области поиска, поэтому при размерности задачи порядка 10 его вычисление может быть выполнено методом перебора.
	\item Формируется многоэкстремальная mixed-integer function 
	\[
	f(y,u) = \left(\varphi(y) + \sum_{j=1}^M{u_j}\right)\left(C + \mu(y,u)\right).
	\]
	По построению $f(y,u)$  будет принимать минимальное значение в точке $(y',b)$.

\end{enumerate}

In the problems generated in our experiments, there were 5 integer  and 3 continuous parameters.
The accuracy of the search was equal to $10^{-2}$.

%All problems of the series have been solved successfully and for solving the problems based on the \textit{Simple} class 11988 trials in average were required whereas for solving the problems based on the \textit{Hard} class 24750 trials were required.

\section{Conclusion}

Заключение

%\textbf{Acknowledgments}. This study was supported by the Russian Science Foundation, project No 16-11-10150.

@@

\bibitem{Gaviano} Gaviano, M., Kvasov, D.E, Lera, D., and Sergeyev, Ya.D.: Software for generation of classes of test functions with known local and global minima for global optimization. ACM Transactions on Mathematical Software \textbf{29}(4), 469--480 (2003)
